### 2020-10-24 Thesis-Chapter3-Arguments.tex
Write incrementalism section

Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to classical problem solving
methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the sweeping
application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
twentieth century, believers in the power of reason thought this progress would march on in the realm of public
planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:

@@

This section describes the two common current policymaking approaches, the classical analytical method and ``muddling
through.''

@@

The classical analytical method \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
\cite{rittel_dilemmas_1973}, ``traditional policy analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking''
\cite{commission_tackling_2018}) has already been introduced. It is the reason-based method that functions by setting
goals, identifying problems, evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct
errors. \myfig{fig-classical-method} illustrates the approach in the context of encryption and \ac{EA}. Per its name, it
has a linear flow from problem to solution. The refinement step reflects the fact that this method tracks outcomes and
tweaks solutions over time; however, refinement is always in-kind---it represents a reinforcement, as opposed to a
reassessment, of the chosen solution.

@@

The shortcomings of the classical analytical method were discussed in \mysec{wicked-characteristics}. It fails due to
disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.

\subsection{Muddling Through}

``Muddling through'' represents an alternative policy making strategy.

See \myfig{fig-muddling-through}.

\begin{figure}[h]
  \centering\CaptionFontSize
  \includegraphics[width=0.55\linewidth]{dfds/build/OODA-muddling.png}
  \caption[The ``Muddling Through'' Method]{The ``Muddling Through'' Method}
  \label{fig-muddling-through}
\end{figure}

% Mention hermeneutics and complex adaptive systems?
% As feeley_judicial_2000 points out, muddling through is too focused on, or even consists entirely of, the
% implementation step. Because of the long tail effects of any implementation in a wicked scenario, and especially in
% the context of encryption, mere muddling through is too risky. Therefore I suggest iteration. Classical methods, but
% applied in a cyclical manner.

>>>>

Rittel's categorization of \acp{wicked-prob} grew out of frustration with the problems' resistance to classical problem
solving methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the
sweeping application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
twentieth century, believers in the power of reason thought this progress would march on in the realm of public
planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:

@@

This section describes two common policymaking approaches, the \ac{classical-method} and \ac{incrementalism}.

@@

The \ac{classical-method} \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
\cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_science_1959}, ``traditional policy
analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking'' \cite{commission_tackling_2018}) has already been
introduced. It is the reason-based method that functions by setting goals, identifying problems, evaluating
alternatives, implementing solutions, and analyzing outcomes in order to correct errors. \myfig{fig-classical-method}
illustrates the approach in the context of encryption and \ac{EA}. Per its name, it has a linear flow from problem to
solution. The refinement step reflects the fact that this method tracks outcomes and tweaks solutions over time;
however, refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
solution.

@@

The shortcomings of the \ac{classical-method} were discussed in \mysec{wicked-characteristics}. It fails due to
disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.

\subsection{Incrementalism}

\Ac{incrementalism} is an intuitive and iterative approach posed as an alternative to the \ac{classical-method}.
\Ac{incrementalism} as a policymaking strategy is often referred to as ``\ac{muddling-through}'' after political
scientist Charles Lindblom's classic paper defining and defending the approach \cite{lindblom_muddling_1959}. Written
before Rittel developed the idea of ``\acp{wicked-prob},'' Lindblom nonetheless identified many of the same shortcomings
of the analytical method and sought to formalize the process policymakers were already often using.

Lindblom's ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis. The
alternatives selected for comparison are defined relative to the status quo and are close enough to one another to be
able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
impossible to impose in a democratic system \cite{lindblom_muddling_1959}. \myfig{fig-muddling-through} illustrates this
method.

%Footnote: see Lindblom's bracing quote: ``Party behavior is in turn rooted in public attitudes, and political theorists
%cannot conceive of democracy's surviving in the United States in the absence of fundamental agreement on potentially
%disruptive issues, with consequent limitation of policy debates to relatively small differences in policy.''

\begin{figure}[h]
  \centering\CaptionFontSize
  \includegraphics[width=0.55\linewidth]{dfds/build/OODA-muddling.png}
  \caption[The Incrementalist Method]{The Incrementalist Method}
  \label{fig-muddling-through}
\end{figure}

\Ac{incrementalism} has several advantages over the \ac{classical-method} for handling \acp{wicked-prob}. It is rooted
in realism about the limits of rational analysis. It accepts that the problem will not be conclusively solved,
emphasizing iteration (``Policy is not made once and for all; it is made and re-made endlessly. Policy-making is a
process of successive approximation to some desired objectives in which what is desired itself continues to change under
reconsideration.'' \cite{lindblom_muddling_1959}). Most importantly, it eschews the linear
problem-definition--alternatives-analysis steps out of respect that this is, indeed, impossible. Lindblom argues at
length that policy ends and means are interlinked, eventually concluding:

\begin{displayquote}
As to whether the attempt to clarify objectives in advance of policy selection is more or less rational than the close
intertwining of marginal evaluation and empirical analysis, the principal difference established is that for complex
[i.e., wicked] problems the first is impossible and irrelevant, and the second is both possible and relevant. The second
is possible because the administrator need not try to analyze any values except the values by which alternative policies
differ and need not be concerned with them except as they differ marginally. His need for information on values or
objectives is drastically reduced as compared with the root [i.e., classical analytic] method; and his capacity for
grasping, comprehending, and relating values to one another is not strained beyond the breaking point.
\cite{lindblom_muddling_1959}
\end{displayquote}

Despite its strengths, \ac{incrementalism} has its weaknesses at addressing \acp{wicked-prob}. Its main weakness is its
lack of high level strategic analysis. It is incapable of drastic change, which is sometimes necessary. By Lindblom's
admission, it lacks a safeguard for consideration of all relevant values and may ``overlook excellent policies for no
other reason than that they are not suggested by the chain of successive policy steps leading up to the present''
\cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to the next can result in messy policy
that none are particularly fond of ("As Lindblom's sobriquet suggests, it often led to a considerable muddle''
\cite{feeley_judicial_2000}). Consider application of the \ac{CFAA} as an example of this kind of muddle
\cite{wolff_computer_2016}.

Unfortunately, one cannot simply expand the strategic view of the incrementalist method by allowing it to look further
into the past than the present status quo. Simplifying analysis by limiting it to marginal differences to a given
baseline is at the heart of \ac{incrementalism}. If one tries to be a ``strategic incrementalist'' by looking into the
past, they still have to choose a baseline from which to perform analysis. This approach is susceptible to two
weaknesses in baseline-based reasoning that Rozenshtein described in his analysis of \ac{EA} as a \ac{wicked-prob}
\cite{rozenshtein_wicked_2018}. First, the choice of baseline is arbitrary, yet heavily colors analysis:

\begin{displayquote}
Both the government and its critics have operated from the status-quo baseline, though from opposite directions. For the
government, the relevant baseline is recent history---specifically, right before companies like Apple and WhatsApp
encrypted their products. From this baseline, the government's ability to surveil has diminished. For critics of
government surveillance, the relevant baseline is the pre-digital age, before smartphones and social media vastly
expanded the government's surveillance capabilities. From this baseline, the technological changes underlying the
``going dark'' problem are mere blips on the otherwise rocketing growth of the surveillance state.
\cite{rozenshtein_wicked_2018}
\end{displayquote}

Second, unlike legal baselines, policy baselines do not carry normative force. Constant changes in the underlying
situation mean that even optimal policy in the past is not necessarily desirable in the present. Due to these
weaknesses, applying incrementalist methods at the strategic level does not work.

A final weakness in \ac{incrementalism} is its assumption of basic agreement and political stability. The method works
by limiting analysis to marginal comparisons of broadly similar and familiar proposals. Proposals that differ widely
from one another or the status quo are considered irrelevant because the debating parties both share the same general
goals and lack the ability to unilaterally impose their will. Rather a symptom of present circumstances (see
\mysec{sec-history-current}) than an inherent weakness in the incrementalist approach, both of these assumptions are
incorrect.

### 2020-10-24 Thesis-Glossary.tex
Make more writing and editing improvements
Add OODA loop diagram and notes; update outline
Write incrementalism section

\newterm{key-escrow}{key escrow}{
    An approach to \ac{EA} that relies on storing additional copies of the encryption key or storing information that
    can be used to derive additional additional copies
}
% End terms

@@

\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
\newabbreviation{SAFEAct}{SAFE Act}{Security and Freedom Through Encryption Act}
\newabbreviation{PROCODEAct}{Pro-C0DE Act}{Promotion of Commerce On-Line in the Digital Era Act}
\newabbreviation{ENCRYPTAct}{ENCRYPT Act}{Ensuring National Constitutional Rights for Your Private Telecommunications Act}
\newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}

>>>>

\newterm{key-escrow}{key escrow}{
    An approach to \ac{EA} that relies on storing additional copies of the encryption key or storing information that
    can be used to derive additional additional copies
}

\newterm{classical-method}{classical analytical method}{
    A policy making strategy that emphasizes rationalism and functions by setting goals, identifying problems,
    evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct errors
}

\newterm{incrementalism}{incrementalism}{
    A policy making strategy that functions by taking successive steps based on the status quo and chosen through
    comparative analysis; also known as ``muddling through''
}

\newterm{muddling-through}{muddling through}{
    See ``incrementalism''
}
% End terms

@@

\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
\newabbreviation{SAFEAct}{SAFE Act}{Security and Freedom Through Encryption Act}
\newabbreviation{PROCODEAct}{Pro-C0DE Act}{Promotion of Commerce On-Line in the Digital Era Act}
\newabbreviation{ENCRYPTAct}{ENCRYPT Act}{Ensuring National Constitutional Rights for Your Private Telecommunications Act}
\newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
\newabbreviation{LEADAct}{LEAD Act}{Lawful Access to Encrypted Data Act}
\newabbreviation{OODAloop}{OODA loop}{\ii{Observe-Orient-Decide-Act} loop}
